---
title: "Math 300 Lesson 32 Notes"
subtitle: "Case Study"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Conduct and interpret a hypothesis test for the difference of two means.      

2. Calculate and interpret a confidence interval for the difference of two means.

## Reading 

[Chapter 9.5](https://moderndive.com/9-hypothesis-testing.html#ht-case-study)

## Lesson 

Work through the learning checks LC 9.9 - LC 9.15.

+ This is a two-sided test. The p-value calculation is a little different. 

+ Interpret the p-value in context of the problem.

***

### Libraries 

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(infer)
library(moderndive)
library(ggplot2movies)
```




### LC 9.9 (Objective 1)

**(LC 9.9)** Conduct the same analysis comparing action movies versus romantic movies using the median rating instead of the mean rating. What was different and what was the same? 

**Solution**:  

```{r}
# Complete the code
set.seed(2511)
# In calculate() step replace "diff in means" with "diff in medians"
# null_distribution_movies_median <- __________ %>% 
#   specify(formula = rating ~ ________) %>% 
#   hypothesize(null = "______________") %>% 
#   generate(reps = 1000, type = "___________") %>% 
#   calculate(stat = "_______________", order = c("Action", "Romance"))
```


```{r}
# Complete the code
# compute observed "diff in medians"
# obs_diff_medians <- ____________ %>% 
#   specify(formula = rating ~ ________) %>% 
#   calculate(stat = "____________", order = c("Action", "Romance"))
# obs_diff_medians
```


```{r}
# Complete the code
# Visualize p-value. Observing this difference in medians under H0
# is very unlikely! Suggesting H0 is false, similarly to when we used
# "diff in means" as the test statistic. 
# visualize(null_distribution_movies_median, bins = 10) + 
#   shade_p_value(obs_stat = obs_diff_medians, direction = "__________")
```


```{r}
# Complete the code
# p-value is very small, just like when we used "diff in means"
# as the test statistic. 
# null_distribution_movies_median %>% 
#   get_p_value(obs_stat = obs_diff_medians, direction = "__________")
```


- Confidence interval

```{r}
# Complete the code
set.seed(2511)
# In calculate() step replace "diff in means" with "diff in medians"
# boot_distribution_movies_median <- movies_sample %>% 
#   specify(formula = rating ~ __________) %>% 
#   generate(reps = 1000, type = "_________") %>% 
#   calculate(stat = "____________", order = c("Action", "Romance"))
```


```{r}
#visualize(boot_distribution_movies_median, bins = 10) 
```


```{r}
# Complete the code
#boot_distribution_movies_median %>% 
#  get_ci(level=_______,type="______________")
```



### LC 9.10 (Objective 1)

**(LC 9.10)** What conclusions can you make from viewing the faceted histogram looking at `rating` versus `genre` that you couldn't see when looking at the boxplot?

**Solution**:  

```{r}
# Complete the code
# ggplot(data = movies_sample, aes(x = __________)) +
#   geom_histogram(bins=8) +
#   facet_wrap(~_______)+
#   labs(x = "IMDb rating")
```




### LC 9.11 (Objective 1)

**(LC 9.11)** Describe in a paragraph how we used Allen Downey's diagram to conclude if a statistical difference existed between mean movie ratings for action and romance movies.

**Solution**:  



### LC 9.12 (Objective 1)

**(LC 9.12)** Why are we relatively confident that the distributions of the sample ratings will be good approximations of the population distributions of ratings for the two genres?

**Solution**:  



### LC 9.13 (Objective 1)

**(LC 9.13)** Using the definition of $p$-value, write in words what the $p$-value represents for the hypothesis test comparing the mean rating of romance to action movies.

**Solution**:  



### LC 9.14 (Objective 1)

**(LC 9.14)** What is the value of the $p$-value for the hypothesis test comparing the mean rating of romance to action movies?

**Solution**: 



### LC 9.15 (Not testable)

**(LC 9.15)** Test your data wrangling knowledge and EDA skills:

- Use `dplyr` and `tidyr` to create the necessary data frame focused on only action and romance movies (but not both) from the `movies` data frame in the `ggplot2movies` package.
- Make a boxplot and a faceted histogram of this population data comparing ratings of action and romance movies from IMDb.
- Discuss how these plots compare to the similar plots produced for the `movies_sample` data.

**Solution**: 

- Use `dplyr` and `tidyr` to create the necessary data frame focused on only action and romance movies (but not both) from the `movies` data frame in the `ggplot2movies` package. 

```{r}
action_romance <- movies %>%
  select(title,year,rating,votes,Action,Romance) %>%
# Get rid of movies that are both
  filter(!(Action == 1 & Romance == 1)) %>%
  filter(Action == 1 | Romance == 1) %>%
   mutate(genre = case_when(
    Action == 1 ~ "Action",
    Romance == 1 ~ "Romance",
    TRUE ~ "Neither"
  )) %>%
  select(-Action,-Romance)
```

Summary stats

```{r}
# Complete code
# action_romance %>% 
#   group_by(_______) %>% 
#   summarize(n = n(), mean_rating = mean(_______), std_dev = sd(_______))
```

- Make a boxplot and a faceted histogram of this population data comparing ratings of action and romance movies from IMDb.


```{r}
# Complete code
# ggplot(data = action_romance, aes(x = __________)) +
#   geom_histogram(bins=8) +
#   facet_wrap(~__________)+
#   labs(x = "IMDb rating")
```

```{r}
# Complete code
# ggplot(data = action_romance, aes(x = ________)) +
#   geom_density() +
#   facet_wrap(~________)+
#   labs(x = "IMDb rating") +
#   theme_classic()
```

```{r}
# Complete code
# ggplot(data = action_romance, aes(x = _________,y=______________)) +
#   geom_boxplot() +
#   labs(y = "IMDb rating") +
#   theme_classic()
```



***

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `moderndive` package version: `r packageVersion("moderndive")`
  * `infer` package version: `r packageVersion("infer")`
  * `ggplot2movies` package version: `r packageVersion("ggplot2movies")`

