---
title: "Math 300 Lesson 34 Notes"
subtitle: "Interpreting Regression Tables"
author: "YOUR NAME HEREr"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1.  Build and interpret a regression table and use correct terminology.

2. Write the hypotheses from the regression table.

3. Use the p-value from a regression table to make a decision about the relationship between the explanatory and response variables.  


## Reading

[Chapter 10 - 10.2](https://moderndive.com/10-inference-for-regression.html)

## Lesson

*Remember that you will be running this more like a lab than a lecture. You want them using `R` and answering questions. Have them open the notes rmd and work through it together.*

There are no learning checks for this section.

-   This is a chance to review regression and also to conduct a hypothesis test.

-   Focus on the interpretation of the results.

------------------------------------------------------------------------

### Libraries

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(infer)
library(moderndive)
```

## Problem

We will now use the inference tools we have developed in a regression settings. In this lesson we will review and spend some time on interpreting regression output.

### Data

Let's get the data again.

```{r}
evals_ch5 <- evals %>%
  select(ID, score, bty_avg, age)
glimpse(evals_ch5)
```

### EDA

The scatterplot revealed what appeared to be a slight positive association between the two variables. But is this just to the randomness associated with sampling?

```{r message=FALSE,warning=FALSE}
# Complete the code
# ggplot(evals_ch5, 
#        aes(x = _________, y = _________)) +
#   geom_point() +
#   labs(x = "Beauty Score", 
#        y = "Teaching Score",
#        title = "Relationship between teaching and beauty scores") +  
#   geom_smooth(method = "lm", se = FALSE) +
#   theme_classic()
```
### Regression Model

Let's build the linear regression model.

```{r}
# Complete the code
# Fit regression model:
# score_model <- lm(________ ~ ________________, data = evals_ch5)
```


```{r}
# Complete the code
# Get regression table:
# get_regression_table(______________)
```

- Some review:

> For every increase of one unit in “beauty” rating, there is an associated increase, on average, of 0.067 units of evaluation score.

The hypothesis test for the slope is:

$$
H_O: \beta_1 = 0 
$$

$$
H_A: \beta_1 \neq 0
$$

> A p-value is the probability of obtaining a test statistic just as extreme or more extreme than the observed test statistic assuming the null hypothesis  $H_O4  is true.

### Conclusion 

**Answer:**

## Your turn

In the `moderndive` package is a dataset on electric vehicle charging seesions for a workplace charging program.

We are interested in there is an association between the total energy used at a station, `kwh_total`, and the amount paid for a charging session, `dollars`. We are only interested those workers that paid more than $0.50 for a session. Repeat the analysis from the book on this dataset. 

### Data

```{r}
ev_chp10 <- ev_charging %>%
  select(kwh_total,dollars) %>%
  filter(dollars>0.5)
```

```{r}
glimpse(ev_chp10)
```







### EDA

Scatterplot 

```{r message=FALSE,warning=FALSE}
ggplot(ev_chp10, 
       aes(y = kwh_total, x = dollars)) +
  geom_point() +
  labs(x = "Money Paid", 
       y = "Total Energy",
       title = "Relationship between money paid for charging session and total energy used") +  
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic()
``` 

### Regression Model

Let's build the linear regression model.

```{r}
# Complete the code
# Fit regression model:

```


```{r}
# Complete the code
# Get regression table:

```


### Conclusion 

**Answer:**


***

## Documenting software

-   File creation date: `r Sys.Date()`
-   `r R.version.string`
-   `tidyverse` package version: `r packageVersion("tidyverse")`
-   `moderndive` package version: `r packageVersion("moderndive")`
-   `infer` package version: `r packageVersion("infer")`

