---
title: "Math 300 Lesson 31 Notes"
subtitle: "Interpreting Hypothesis Tests"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Correctly use terminology and notation of interpreting hypothesis tests.     



## Reading 

[Chapter 9.4](https://moderndive.com/9-hypothesis-testing.html#ht-interpretation)


## Lesson 

Work through the learning checks LC 9.5 - LC 9.8.

+ This section is conceptually difficult. We are setting up a contradiction but it is based on empirical data. Thus the language becomes somewhat confusing. The criminal trial analogy is a good reference to put the ideas into context. There is a large number of terms.  

+ The fail to reject idea is tricky since people want to "accept the null hypothesis". Remember there is a difference between "not guilty" and "innocent" in a trial. 

+ This framework helps:  
a - The defendant is truly either “innocent” or “guilty.” Two possible outcomes of the test.  
b - The defendant is presumed “innocent until proven guilty.” The null hypothesis.   
c - The defendant is found guilty only if there is strong evidence that the defendant is guilty. The phrase “beyond a reasonable doubt” is often used as a guideline for determining a cutoff for when enough evidence exists to find the defendant guilty. "Beyond a reasonable doubt" is the level of significance $\alpha$.  
d - The defendant is found to be either “not guilty” or “guilty” in the ultimate verdict. This is "reject" or "fail to reject". 

+ Write conclusion in the context of the problem. For example "We found enough evidence in this data to suggest that there was gender discrimination at play in bank promotions."  

+ The names **Type I** and **Type II** are not informative. False positive and false negative make more sense in the context that a positive is rejecting the null hypothesis. 

***

### Libraries 

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(infer)
library(moderndive)
library(nycflights13)
library(ggplot2movies)
```


```{r, include=FALSE, purl=FALSE}
chap <- 9
lc <- 4
```

### LC 9.5 (Objective 1)


**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**  What is wrong about saying, "The defendant is innocent." based on the US system of criminal trials?

**Solution**:  



### LC 9.6 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**  What is the purpose of hypothesis testing?

**Solution**:  



### LC 9.7 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**  What are some flaws with hypothesis testing?  How could we alleviate them?

**Solution**:  



### LC 9.8 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`**  Consider two $\alpha$ significance levels of 0.1 and 0.01. Of the two, which would lead to a more *liberal* hypothesis testing procedure? In other words, one that will, all things being equal, lead to more rejections of the null hypothesis $H_0$.

**Solution**:  





## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `moderndive` package version: `r packageVersion("moderndive")`
  * `infer` package version: `r packageVersion("infer")`

