---
title: "Math 300 NTI Lesson 11"
subtitle: "Simple Linear Regression - Continuous Predictor"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Use the `skimr` package to summarize multiple numerical variables in a data frame.

2. Build a scatterplot to describe the relationship between two continuous, numerical variables; use `geom_smooth()` to visualize the best fit line. 

3. Fit a linear regression model between two variables using the `lm()` function and interpret the output. This includes the interpretation of slope and the use of *association* and not *causation*.    

4. Generate a table of observations, fitted values, and residuals from a linear regression object.  

## Reading 

[Chapter 5 - 5.1](https://moderndive.com/5-regression.html)

## Lesson 

*Remember that you will be running this more like a lab than a lecture. You want them using `R` and answering questions. Have them open the notes rmd and work through it together.*

Work through the learning checks LC5.1 - LC5.3.

+ Regression can be used for explanatory and predictive purposes. It falls on that line between traditional statistics/econometrics and machine learning. In this course we focus on its more traditional use to interpret the relationship between predictors and a response. Math 378 is our machine learning course and expands on linear regression in this framework.   

+ Note the many different terms for `x` and `y` in regression. These names come from different fields. For example, `y` is called the response, dependent variable, outcome, and output. Meanwhile, `x` is called input, predictor, independent variable, and explanatory variable. Also point out that in linear regression, `y` is numerical while `x` can be numerical or categorical.

+ We are using new packages. The `tidyverse` package is a wrapper and actually loads `readr`, `dplyr`, `ggplot2`, and `tidyr`.  

+ In the reading, the authors setup the problem with instructor teaching score as the response and beauty score as the explanatory variable. What is the research question? 

+ The reading introduces tilde `~` as a formula. You might want to talk about this as we use it in LC 5.1. 

+ The interpretation of the slope has the key phrase **average**. For a one unit change in `x`, the average value of `y` changes by the value of the slope. 


### Setup

```{r, include=FALSE, purl=FALSE}
chap <- 5
lc <- 0
```

```{r message=FALSE}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
``` 

*Create the data needed for the exercises.*

```{r}
evals_ch5 <- evals %>%
  select(ID, score, bty_avg, age)
```

Let's look at 5 random rows of data.

```{r}
set.seed(1234)
evals_ch5 %>%
  sample_n(size = 5) 
```

### LC 5.1 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Refer to the example in section 5.1.1. Conduct a new exploratory data analysis with the same outcome variable $y$ being `score` but with `age` as the new explanatory variable $x$. Remember, this involves three things:

- Looking at the raw data values.
- Computing summary statistics.
- Creating data visualizations.

What can you say about the relationship between age and teaching scores based on this exploration?

**Solution**: 

- Looking at the raw data values: 
```{r}
glimpse(evals_ch5)
```

- Computing summary statistics:
```{r, eval=FALSE}
my_skim<-skim_with(numeric = sfl(hist = NULL))
```


```{r, include=FALSE}
my_skim<-skim_with(numeric = sfl(hist = NULL))
```

```{r}
evals_ch5 %>%
  select(score, age) %>%
  my_skim() %>%
  print()
```  


(Note that for formatting purposes, the inline histogram that is usually printed with skim() has been removed.)

- Bivariate summary:

```{r}
evals_ch5 %>% 
  get_correlation(formula = score ~ age)
```


- Creating data visualizations:
```{r message=FALSE}
ggplot(evals_ch5, aes(x = age, y = score)) +
  geom_jitter(alpha=0.5) +
  labs(
    x = "Age", y = "Teaching Score",
    title = "Scatterplot of relationship of teaching score and age") +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic()
```

Based on the scatterplot, there does not appear to be a relationship between age and teaching score. If anything, there might be a slight negative linear trend. That is, as age increases, the **average** teaching score decreases slightly. Even thought the correlation coefficient is negative, it is small in absolute value and thus there may be no relationship between the variables.  

### LC 5.2 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Fit a new simple linear regression using `lm(score ~ age, data = evals_ch5)` where `age` is the new explanatory variable `x`. Get information about the "best-fitting" line from the regression table by applying the `get_regression_table()` function. How do the regression results match up with the results from your earlier exploratory data analysis?

**Solution**:  

```{r}
# Fit regression model:
score_age_model <- lm(score ~ age, data = evals_ch5)
```


```{r}
# Get regression table:
get_regression_table(score_age_model)
```


$$
\begin{aligned}
\widehat{y} &= b_0 + b_1 \cdot x\\
\widehat{\text{score}} &= b_0 + b_{\text{age}} \cdot\text{age}\\
&= 4.462 - 0.006\cdot\text{age}
\end{aligned}
$$

For every increase of 1 year in `age`, there is an *associated* decrease of 0.006 units of the **average** teaching `score`. It matches with the results from our earlier exploratory data analysis. 

### LC 5.3 (Objective 3)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Generate a data frame of the residuals of the model where you used `age` as the explanatory $x$ variable.

**Solution**: 

```{r}
score_age_regression_points <- get_regression_points(score_age_model)
```


```{r}
head(score_age_regression_points)
```



## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `skimr` package version: `r packageVersion("skimr")`
  * `gapminder` package version: `r packageVersion("gapminder")`  
  * `moderndive` package version: `r packageVersion("moderndive")`
