
---
title: "Problem Set 01 Solution"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document: default
  html_document:
    theme: lumen
---

```{r message = F}
library(dplyr)
library(ggplot2)
library(readr)
```


```{r, message = F}
nc <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTm2WZwNBoQdZhMgot7urbtu8eG7tzAq-60ZJsQ_nupykCAcW0OXebVpHksPWyR4x8xJTVQ8KAulAFS/pub?gid=202410847&single=true&output=csv")
```


```{r eval = F}
glimpse(nc)
```

## Exercise 1

Looking at the output of the `glimpse()` function, has does `R` classify the variable `mature`? Same question for the variable `gained`. (Answer with text)

**Answer: mature is listed as <chr> (character), and visits is an double precision <dbl>**




## Exercise 2

Make a graph showing a mother's age `mage` on the x axis and the variable `weeks` on the y axis.  Include axis labels with measurement units, and a title. (`R` code and output)

**Answer:**

```{r}
ggplot(data = nc, aes(x = mage, y = weeks))+ 
  geom_point() + 
  labs(x = "Mother's Age (years)", y = "Pregnancy Length (weeks)", 
       title = "Relationship between mother's age and pregnancy duration") +
  theme_classic()
```


## Exercise 3  

Study the code below, and the resulting graphical output. Note that we added a new argument of `color = premie` and `shape = marital` **inside** the `aes`thetic mapping. The variable `premie` indicates whether a birth was early (premie) or went full term and `marital` represents the marital status of the mother. Please answer with text:

**A.** What did adding the argument `alpha = 0.3` accomplish? 
     
**Answer: It makes the data points more transparent so we can see overlapping points.**
     
     
**B.** How many **variables** are now displayed on this plot?  

**Answer: 4 variables are shown**
     
**C.** What appears to (roughly) be the pregnancy length cutoff for classifying a newborn as a "premie"" versus a "full term". 

**Answer: anywhere between 36 and 38 seems reasonable**

```{r fig.height=3, fig.width=5, warning=F}
ggplot(data = nc, aes(x = weeks, y = gained, color = premie, shape = marital))+ 
  geom_point(alpha=0.3) + 
  labs(x = "Pregnancy length (wks)", y = "Maternal weight gain (lbs)") +
  theme_light()
```


## Exercise 4  

Make a new scatterplot that shows a mothers age on the x axis (variable called `mage`) and pregnancy length on the y axis (`weeks`). Color the points on the plot based on the gender of the resulting baby (variable called `gender`). Change the shape based on the marital status of the mother (variable called `marital`). Use `alpha` value of 0.4 and use the classic theme. There should not appear to be any strong relationship between a mother's age and the pregnancy length. (`R` code and output)

**Answer:**

```{r include = F}
ggplot(data = nc, aes(x = mage, y = weeks, color = gender, shape = marital)) + 
  geom_point(alpha = 0.4) + 
  labs(x = "Mother's age (in years)", y = "Pregnancy length (weeks)", 
       title = "Pregnancy length by age, baby gender, and marital status") +
  theme_classic()
```


## Exercise 5  

```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x = mage))+ 
  geom_histogram(binwidth = 1, color = "white", fill = "steelblue") +
  theme_classic()
```

Inspect the histogram of the `mage` variable. Answer each of the following with **text**. 

**A.** The y axis is labeled **count**. What is specifically being counted in this case? Hint: think about what each case is in this data set. 

**Answer: the number of mothers whose age fall into each bin specified on the histogram**
 
    
**B.** What appears to be roughly the average mother's age in years?
    

**Answer: 27 weeks...26, 28, 29, 30 or 31 also acceptable**
    


## Exercise 6  

Make a histogram of the birth `weight` of newborns (which is in lbs), including a title and axis labels. (`R` code and output for answer)

```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x =  weight)) +
  geom_histogram(binwidth = 1, color = "white", fill = "sienna") + 
  labs(x = "Newborn weight (lbs)", y = "Count", title = "Distribution of newborn birth weights")
  
```

## Exercise 7  

Make a histogram of newborn birth `weight` split by `habit`, the smoking status of the mother. Set the binwidth to 0.5. There are many fewer mothers who smoke. Do nonsmokers appear to have babies with a slightly larger median birth weight? (Text and `R` code and output for answer)

**Answer:** 

```{r}
ggplot(data = nc, aes(x = weight)) +
  geom_histogram(binwidth = 0.4, color = "white", fill = "steelblue") +
  facet_wrap(~ habit, ncol = 1)
```


**Answer: The babies of nonsmokers tend to have a slightly higher median birth weight**


## Exercise 8  

Make a boxplot of the weight `gained` by mothers, split by the marital status of the mothers (`marital`).  Include axis labels and a title on your plot. Is the variation in weight gained during pregnancy larger for married or single mothers? (Text and `R` code and output)

**Answer:** 

```{r}
ggplot(data = nc, aes(x = marital, y = gained)) + 
  geom_boxplot(fill = "steelblue") + 
  labs(x = "Marital status of mother", y = "Pregnancy weight gain (lbs)", 
       title = "Gravida weight gain for single and married mothers")
```

**Answer: The variation is slighly greater for married mothers**

## Exercise 9  

Make a boxplot of pregnancy duration in `weeks` by smoking `habit`. Are there more outliers in the duration of pregnancy for smokers or non-smokers? (Text and `R` code and output for answer)

**Answer:**

```{r}
ggplot(data = nc, aes(x = habit, y = weeks)) + 
  geom_boxplot(fill = "steelblue")
```

**Answer: The nonsmokers have more outliers, premature babies. We are tempted to assign a cause such as the mothers are more concerned with their pregnancies and would not risk smoking. But this is only a conjecture that we have to be careful in not making.**

## Exercise 10  

Make a side-by-side barplot of marital status `marital` and gender of the baby `gender`. Are female babies more common in both married and single mothers? (Text and `R` code and output for answer)

**Answer:**  

```{r}
ggplot(data = nc, aes(x = marital, fill = gender)) + 
  geom_bar(position = "dodge")
```

**Answer: Female babies are more common in both married and single mothers.**

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `dplyr` package version: `r packageVersion("dplyr")`
  * `ggplot2` package version: `r packageVersion("ggplot2")`
  * `readr` package version: `r packageVersion("readr")`


