---
title: "Math 300 Lesson 27 Notes"
subtitle: "Case Study"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Calculate, both percentile and standard error method, confidence intervals for a difference of two proportions.   

2. Calculate a normal-based, theory-based, confidence interval for a single proportion or mean.

3. Correctly use all terms and notation to include standard error, margin of error, sampling distribution, and bootstrap distribution.  

4. Compare and contrast sampling distribution and bootstrap distribution.  

## Reading 

[Chapter 8.6](https://moderndive.com/8-confidence-intervals.html#case-study-two-prop-ci)


## Lesson 

There are no learning checks for this lesson. 

+ The case study brings all the ideas together while introducing a two-sample test. 

+ The bootstrap sample needs to be the same as the original sample. What is not clear is if they sampled such that there will be 16 in the control group and 34 in the seed group. The `infer` package does not allow us to sample within groups so we sample with replacement from the entire sample.  


***

### Libraries 

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(moderndive)
library(infer)
```


### Yawning Problem

The reading discusses the research question "Is yawning contagious?". Let's summarize the data in two different ways.

```{r}
mythbusters_yawn %>% 
  group_by(group, yawn) %>% 
  summarize(count = n()) %>%
  mutate(prop = count / sum(count))
```

- What does this summary tell the reader? This is finding proportions conditioned on groups.

**Answer:** 

Now another summary is as follows:

```{r}
# Complete the code
# mythbusters_yawn %>% 
#   group_by(group, _____) %>% 
#   summarize(count = ____) %>%
#   ungroup() %>%
#   mutate(prop = count / sum(count))
```

- What does this summary tell us about the data?  

**Answer:** 

- Which summary is better for the research question?

**Answer:** 


Let's determine how the `infer` package sample the data.


```{r}
# Complete the code
# set.seed(77)
# mythbusters_yawn %>% 
#   specify(formula = ______ ~ _______,success="______") %>% 
#   generate(reps = 1, type = "__________") %>% 
#   group_by(group, yawn) %>%
#   summarize(count = n()) %>%
#   mutate(prop = count / sum(count))
```

The numbers in the control and seed group changed. The sample is with replacement from the entire data frame and not within groups.

### Bootstrap Sampling Distribution

#### Generate the bootstrap distribution

```{r}
# Complete the code
# bootstrap_distribution_yawning <- mythbusters_yawn %>% 
#   specify(formula = yawn ~ group, success = "_______") %>% 
#   generate(reps = 1000, type = "bootstrap") %>% 
#   calculate(stat = "___________", order = c("_______", "________"))
```


```{r}
#head(bootstrap_distribution_yawning)
```

#### Visualize the results.

```{r}
# Complete the code
#visualize(________________) 
```

- Where is the center of this distribution?

**Answer:** 

```{r}
# Complete the code
# bootstrap_distribution_yawning %>%
#   summarize(ave=mean(___________))
```



#### Confidence Interval

```{r}
# Complete the code
# get_ci(bootstrap_distribution_yawning,type="_________")
```

- Interpret this confidence interval.

**Answer:** 



***

### Theory-based confidence interval

For the data object `bowl_sample_1` find a 95% confidence interval using the bootstrap percentile, bootstrap standard error method, and the theory-based method.

```{r}
head(bowl_sample_1)
```

- Find the bootstrap percentile confidence interval for this problem.

**Answer:**

```{r}
# Complete the code
# bowl_sample_1 %>%
#   summarize(prop=__________(color=="_______"))
```

```{r}
# Complete the code
# set.seed(45678)
# bootstrap_dist<-bowl_sample_1 %>%
#   specify(color~________,success="________") %>%
#   generate(reps=1000,type="bootstrap") %>%
#   calculate(stat="_________")
```

```{r}
# get_ci(bootstrap_dist,type="_________")
```

- Find the bootstrap percentile confidence interval for this problem.

**Answer:**

```{r}
# get_ci(bootstrap_dist,type="________",point_estimate = ______)
```


- Find the theory-based confidence interval for this problem.

**Answer:**

```{r}
# Complete the code
# bowl_sample_1 %>%
#   summarize(prop=mean(color=="_______"),total=n()) %>%
#   mutate(lower_ci=prop-qnorm(________)*sqrt(prop*(1-prop)/total),
#          upper_ci=prop+qnorm(_______)*sqrt(______________))
```

***

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `moderndive` package version: `r packageVersion("moderndive")`
  * `infer` package version: `r packageVersion("infer")`

