---
title: "Math 300 NTI Lesson 25"
subtitle: "Constructing Confidence Intervals"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Construct bootstrap percentile and standard error confidence intervals for a single mean or median using the `infer` package.

## Reading 

[Chapter 8.4](https://moderndive.com/8-confidence-intervals.html#bootstrap-process)


## Lesson 

*Remember that you will be running this more like a lab than a lecture. You want them using `R` and answering questions. Have them open the notes rmd and work through it together.*

Work through the learning check LC 8.5.

+ We will be using the `infer` package to create the bootstrap confidence intervals. Last lesson we used `rep_sample_n()`. The `infer` package will be used for inference in the rest of the course.

+ The `infer` package gives a framework to think about and conduct inference. It makes hypothesis testing and confidence interval construction more structured and puts computational resources at the center versus mathematical tools.

+ the `infer` package uses the verbs `specify()`, `generate()`, `calculate()`, and `visualize()` to complete the construction process.


***

### Libraries 

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(moderndive)
library(infer)
```

```{r, include=FALSE, purl=FALSE}
chap <- 8
lc <- 4
```

### Review

As a review, let's construct the bootstrap distribution of the sample mean for our pennies data. This requires use to use our skills on wrangling data frames.

```{r}
head(pennies_sample)
```


```{r}
set.seed(52249)
bootstrap_dist<-pennies_sample %>%
  rep_sample_n(size = 50, replace = TRUE, reps = 1000) %>%
  group_by(replicate) %>%
  summarize(ave=mean(year))
```

```{r}
bootstrap_dist %>%
  ggplot(aes(x=ave)) +
  geom_histogram(binwidth = 1, color = "black", boundary = 1990, fill = "cyan") +
  theme_classic()
```

```{r}
bootstrap_dist %>%
  summarize(center=mean(ave),
            lower=quantile(ave,probs=0.25),
            upper=quantile(ave,probs=.975))
```

The standard error method.

```{r}
bootstrap_dist %>%
  summarize(center=mean(ave),
            lower=center-sd(ave)*qnorm(.975),
            upper=center+sd(ave)*qnorm(.975))
```

### Steps from the `infer` package

- First specify the response variable, and explanatory variables if present. We like to use the `formula` option.

```{r}
pennies_sample %>%
  specify(formula=year~NULL) %>%
  head()
```

It is similar to using `select()` but note the meta data has also changed.

- Generate replicates

```{r}
pennies_sample %>% 
  specify(formula=year~NULL) %>% 
  generate(reps = 1000, type = "bootstrap") %>%
  head()
```

Compare this code with that using `rep_sample_n()`.

- Find the sample statistic for each replicate.

```{r}
pennies_sample %>% 
  specify(formula=year~NULL) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean") %>%
  head()
```

- Using results of bootstrap distribution
 
```{r}
boot_dist_mean <- pennies_sample %>% 
  specify(formula=year~NULL) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

```{r}
(percentile_ci <- boot_dist_mean %>% 
  get_confidence_interval(level = 0.95, type = "percentile"))
```


```{r}
# Visualize the results
visualize(boot_dist_mean) +
  shade_confidence_interval(endpoints = percentile_ci)
```

Or if we want the standard error method.

```{r}
(mean_pennies <- pennies_sample %>%
  summarize(ave=mean(year)) %>%
  pull())
```


```{r}
(standard_error_ci <- boot_dist_mean %>% 
  get_confidence_interval(type = "se", point_estimate = mean_pennies))
```


### LC 8.5 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Construct a 95% confidence interval for the *median* year of minting of *all* US pennies? Use the percentile method and, if appropriate, then use the standard-error method.

**Solution**: 

Using the percentile method:

```{r}
set.seed(539)
bootstrap_distribution <- pennies_sample %>%
  specify(formula = year ~ NULL) %>%
  generate(reps = 1000, type="bootstrap") %>%
  calculate(stat = "median")
```


```{r}
percentile_ci <- bootstrap_distribution %>%
  get_confidence_interval(level = 0.95, type = "percentile")
```


```{r}
percentile_ci
```

The standard-error method is not appropriate, because the bootstrap distribution is not bell-shaped:

```{r}
visualize(bootstrap_distribution)
```

Let's visualize the interval.

```{r}
visualize(bootstrap_distribution) + 
  shade_confidence_interval(endpoints = percentile_ci) +
  theme_classic()
```


***

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `moderndive` package version: `r packageVersion("moderndive")`
  * `infer` package version: `r packageVersion("infer")`

