---
title: "Math 300 NTI Lesson 19"
subtitle: "Sampling"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Explain the need for randomization in sampling. 

2. Explain the impact of sample size on the sample distribution and number of replications on the sampling procedure.  

3. Using a sampling distribution, make decisions about the population. In other words, understanding the effect of sampling variation on our estimates.

## Reading 

[Chapter 7 - 7.2](https://moderndive.com/7-sampling.html)

## Lesson 

*Remember that you will be running this more like a lab than a lecture. You want them using `R` and answering questions. Have them open the notes rmd and work through it together.*

Work through the learning checks LC7.1 - 7.7.

+ Learning checks 7.3 - 7.5 are not in the book but are in the notes for students. We are not sure why they were dropped. 

+ Our solutions may be different from those in the back of the book. Use these as discussion points.

+ This idea of an estimate from a sample to make a decision about the population is difficult. In 7.1, there is repeated sampling but in practice there is only one sample. Thus the idea of a sampling distribution tends to be difficult for students. Spend some time on the `tactile_prop_red` data and the analysis done in the book.  

### Setup

```{r, include=FALSE, purl=FALSE}
chap <- 7
lc <- 0
```

```{r message=FALSE}
library(tidyverse)
library(moderndive)
``` 

Here is the data from the reading and the histogram. 

```{r}
head(tactile_prop_red)
```

- What is each row in the data? 

Take a moment to really emphasize this. Each row represents one random sample of size 50 from the bucket. In reality, we won't have many such samples, but just one. The purpose of this exercise is to demonstrate how that sampling distribution behaves. 

- Is this data tidy?  

```{r}
ggplot(tactile_prop_red, aes(x = prop_red)) +
  geom_histogram(binwidth = 0.05, boundary = 0.4, color = "white") +
  labs(x = "Proportion of 50 balls that were red", 
       title = "Distribution of 33 proportions red") +
  theme_classic()
```

- What do you think is the true proportion of red balls in the bin?  


### LC 7.1 (Objective 1)


**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Why was it important to mix the bowl before we sampled the balls?
  
  **Solution**:  
  
  So that we make sure the sampled balls are randomized and thus reduce the chance that our result is biased. Suppose all the red ball were in one side of the bin and we sampled from that side. It would give us an unreasonably high estimate of the proportion of red balls.  


### LC 7.2 (Objective 1) 

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Why is it that our 33 groups of friends did not all have the same numbers of balls that were red out of 50, and hence different proportions red?
  
  **Solution**:  
  
Because it is a sample and by chance you get different proportions of red balls. With randomization and 33 different pairs, it is highly unlikely that all groups would get the same proportion of red balls. 

### LC 7.3 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Why couldn't we study the effects of sampling variation when we used the virtual shovel only once? Why did we need to take more than one virtual sample (in our case 33 virtual samples)?

**Solution**: 

With one data point, there is no variation. We need to take more than one virtual sample to get a range of proportions.

### LC 7.4 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Why did we not take 1000 "tactile" samples of 50 balls by hand?

**Solution**: 

That would be way too much repeated work. Too expensive. 

### LC 7.5 (Objective 3)

*Code for this problem*

```{r}
# Segment 2: sample size = 50 ------------------------------
# Virtually use shovel 1000 times
set.seed(107)
virtual_samples_50 <- bowl %>% 
  rep_sample_n(size = 50, reps = 1000)
```


```{r}
# Compute resulting 1000 replicates of proportion red
virtual_prop_red_50 <- virtual_samples_50 %>% 
  group_by(replicate) %>% 
  summarize(red = sum(color == "red")) %>% 
  mutate(prop_red = red / 50)
```

```{r}
summary(virtual_prop_red_50)
```


```{r}
#Plot distribution via a histogram
ggplot(virtual_prop_red_50, aes(x = prop_red)) +
  geom_histogram(binwidth = 0.05, boundary = 0.4, color = "white") +
  labs(x = "Proportion of 50 balls that were red", title = "50")  +
  theme_classic()
```


**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Looking at the figure we just created, would you say that sampling 50 balls where 30% of them were red is likely or not? What about sampling 50 balls where 10% of them were red?

**Solution**: 

Using the figure, less than 150 out of the 1000 counts were 30% red. So we would say that sampling 50 balls where 30% of them were red is not very likely. No count was less than 16% red, so sampling 50 balls where 10% of them were red is extremely unlikely. 

### LC 7.6 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** In Figure 7.9, we used shovels to take 1000 samples each, computed the resulting 1000 proportions of the shovel's balls that were red, and then visualized the distribution of these 1000 proportions in a histogram. We did this for shovels with 25, 50, and 100 slots in them. As the size of the shovels increased, the histograms got narrower. In other words, as the size of the shovels increased from 25 to 50 to 100, did the 1000 proportions

- A. vary less,
- B. vary by the same amount, or
- C. vary more?
  
  **Solution**: 
  
  A. As the histograms got narrower, the 1000 proportions varied less. 

### LC 7.7 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** What summary statistic did we use to quantify how much the 1000 proportions red varied?
  
- A. The inter-quartile range
- B. The standard deviation
- C. The range: the largest value minus the smallest.

**Solution**: 
  
  B. The standard deviation is used to quantify how much a set of data varies. 




## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `moderndive` package version: `r packageVersion("moderndive")`
