---
title: "Math 300 NTI Lesson 22"
subtitle: "CLT, Normal Distribution"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Explain the central limit theorem.  

2. For a normal distribution, use `R` to find probabilities and percentiles. 

## Reading 

[Chapter 7.5](https://moderndive.com/7-sampling.html#sampling-conclusion-central-limit-theorem)

[Appendix A.2](https://moderndive.com/A-appendixA.html#appendix-normal-curve)


## Lesson 

*Remember that you will be running this more like a lab than a lecture. You want them using `R` and answering questions. Have them open the notes rmd and work through it together.*

Work through the learning checks LCA.1 - LCA.2.

+ The central limit theorem is difficult. The key idea is that if we create a point estimate that is using summation, then the **distribution of this point estimate** tends to be normal. The larger the sample, the closer the distribution is to being normal. If we are describing a sample mean, then the larger the sample, the smaller the standard error.

+ Use `R` to find probabilities. This implies the use of `pnorm()` and `qnorm()`. You will need to spend time on the use of these functions. 

***

Let's do a simulation.

Suppose we record the time of arrival of 4 friends.

```{r warning=FALSE,message=FALSE}
library(tidyverse)
```


```{r}
set.seed(316)
(time<-round(runif(4,5,6),2))
```
We are generating 4 random numbers in the interval 5 to 6. We record these and find the average.

Take a note to explain `runif(4,5,6)`. We are selecting 4 numbers at random between the values 5 and 6. Using `set.seed(316)` ensures we all end up with the same four numbers in this exercise. 

```{r}
mean(time)
```

This is just one value. What is the distribution of the mean if we were to repeat this? Since this is a simulation, we can do this many, many times and plot the distribution of the mean. Don't worry about the code for doing this.

```{r}
# Repeat this 1000 times
rep_data<-tibble(ave=replicate(1000,mean(round(runif(4,5,6),2))))
```


```{r}
summary(rep_data)
```

```{r message=FALSE,warning=FALSE}
ggplot(rep_data,aes(x=ave)) +
  geom_histogram() +
  xlim(5.1,5.9) +
  theme_classic()
```

Notice how this distribution of is approximately normal. This is a demonstration of the central limit theorem. If we used 10 points per sample instead of 4, the standard error would be smaller. 

```{r}
# Repeat this 1000 times
rep_data2<-tibble(ave=replicate(1000,mean(round(runif(10,5,6),2))))
```


```{r message=FALSE,warning=FALSE}
ggplot(rep_data2,aes(x=ave)) +
  geom_histogram() +
  xlim(5.1,5.9) +
  theme_classic()
```

- Run the code again, but change the sample size to something larger.


```{r, include=FALSE, purl=FALSE}
chap <- "A"
lc <- 0
```

***

### LC A.1 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Say you have a normal distribution with mean $\mu=6$ and standard deviation $\sigma=3$. What proportion of the area under the normal curve is less than 3? Greater than 12? Between 0 and 12?

**Solution:**

* Less than 3: 3 is one standard deviation less than the mean of 6, since $\frac{3 - \mu}{\sigma} = \frac{3-6}{3} = -1$. Thus we compute the area to the left of z = -1 in Figure A.2: 0.15% + 2.35% + 13.5% = 16%.  

```{r}
pnorm(3,mean=6,sd=3)
```


* Greater than 12: 12 is two standard deviations greater than the mean of 6, since $\frac{12 - \mu}{\sigma} = \frac{12-6}{3} = 2$, we compute the area to the right of z = 2 in Figure A.2: 2.35% + 0.15% = 2.5%.

```{r}
pnorm(12,mean=6,sd=3,lower.tail = FALSE)
```


* Between 0 and 12: 0 is two standard deviations less than the mean of 6, since $\frac{0 - \mu}{\sigma} = \frac{0-6}{3} = -2$. Thus we compute the area between z = -2 and z = +2 in Figure A.2: 13.5% + 34% + 34% + 13.5% = 95%.


```{r}
pnorm(12,mean=6,sd=3)-pnorm(0,mean=6,sd=3)
```

### LC A.2 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** (Same normal distribution.) What is the 2.5th percentile of the area under the normal curve? The 97.5th percentile? The 100th percentile?

**Solution:**

* 2.5th percentile: Starting from the left of Figure A.2, since 0.15% + 2.35% = 2.5%, the 2.5th percentile is z = - 2. However, this is in standard units. Thus we need the value in the normal distribution that is two standard deviations lower than the mean: $\mu - 2 \cdot\sigma = 6 - 2 \cdot 3 = 0$.

```{r}
qnorm(.025,mean=6,sd=3)
```
This is not 0 since the figure is an approximation.

* 97.5th percentile: Starting from the left of Figure A.2, since 0.15% + 2.35% + 13.5% + 34% + 34% + 13.5% = 97.5%, the 97.5th percentile is z = +2. However, this is in standard units. Thus we need the value in the normal distribution that is two standard deviations higher than the mean: $\mu + 2 \cdot\sigma = 6 + 2 \cdot 3 = 12$.

```{r}
qnorm(.975,mean=6,sd=3)
```


* 100th percentile: $+\infty$. In other words, 100% of values will be less than positive infinity. 

```{r}
qnorm(1,mean=6,sd=3)
```

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`

