---
title: "Math 300 NTI Lesson 39"
subtitle: "Case Study"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Using all the tools and ideas from the course, complete a study that uses the complete data analysis cycle.

## Reading

[Chapter 11 - 11.2](https://moderndive.com/11-thinking-with-data.html#review)

## Lesson

Complete learning check 11.1.

+ By completing the entire data analysis cycle, we will review many of the ideas from the course. This will help prepare us for the final.  


------------------------------------------------------------------------

### Libraries

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(broom)
library(infer)
library(moderndive)
library(skimr)
```

## Case Study

First let's work through the case study in the reading.

The `house_prices` dataset consists of 21,613 houses and 21 variables describing the sale prices of homes sold between May 2014 and May 2015 in King County, Washington, US (for a full list and description of these variables, see the help file by running ?house_prices in the console). In this case study, we’ll create a multiple regression model where:

The outcome variable $y$ is the sale `price` of houses.
Two explanatory variables:
- A numerical explanatory variable $x_1$: house size `sqft_living` as measured in square feet of living space. Note that 1 square foot is about 0.09 square meters.
- A categorical explanatory variable $x_2$: house `condition`, a categorical variable with five levels where `1` indicates “poor” and `5` indicates “excellent.”

### Exploratory

Let's get the data and explore them.

Recall the three common steps in an exploratory data analysis we introduced in Subsection 5.1.1:

- Looking at the raw data values.  
- Computing summary statistics.  
- Creating data visualizations.  

```{r}
glimpse(house_prices)
```
Now summary statistics.


```{r}
my_skim<-skim_with(numeric = sfl(hist = NULL))
```

```{r}
house_prices %>% 
  select(price, sqft_living, condition) %>% 
  my_skim() %>%
  print()
```  

Let’s now perform the last of the three common steps in an exploratory data analysis: creating data visualizations.

```{r}
# Histogram of house price:
ggplot(house_prices, aes(x = price)) +
  geom_histogram(color = "black", fill = "cyan") +
  labs(x = "price (USD)", title = "House price") +
  theme_classic()
```


```{r}
# Histogram of sqft_living:
ggplot(house_prices, aes(x = sqft_living)) +
  geom_histogram(color = "black", fill = "cyan") +
  labs(x = "living space (square feet)", title = "House size") +
  theme_classic()
```


```{r}
# Density plot of sqft_living:
ggplot(house_prices, aes(x = sqft_living)) +
  geom_density(color = "black", fill = "cyan") +
  labs(x = "living space (square feet)", title = "House size") +
  theme_classic()
```

```{r}
# Barplot of condition:
ggplot(house_prices, aes(x = condition)) +
  geom_bar(color = "black", fill = "cyan") +
  labs(x = "condition", title = "House condition") +
  theme_classic()
```

The distribution of the variables are skewed, so let's transform the variables. Let’s create new log10 transformed versions of the right-skewed variable `price` and `sqft_living` using the `mutate()` function from Section 3.5, but we’ll give the latter the name `log10_size`, which is shorter and easier to understand than the name `log10_sqft_living`.

```{r}
house_prices_reduced <- house_prices %>%
  mutate(
    log10_price = log10(price),
    log10_size = log10(sqft_living)) %>%
  select(log10_price, log10_size, condition)
```

Let's plot the new variables.

```{r}
# Density plot of log10 sqft_living:
ggplot(house_prices_reduced, aes(x = log10_size)) +
  geom_density(color = "black", fill = "cyan") +
  labs(x = "log10 living space (square feet)", title = "House size") +
  theme_classic()
```


```{r}
# Density plot of log10 price:
ggplot(house_prices_reduced, aes(x = log10_price)) +
  geom_density(color = "black", fill = "cyan") +
  labs(x = "log10 price (USD)", title = "House size") +
  theme_classic()
```

These variables seem to be more symmetrical.

We are going to revise our multiple regression model to use our new variables:

The outcome variable $y$ is the sale `log10_price` of houses.
Two explanatory variables:  
- A numerical explanatory variable $x_1$: house size `log10_size` as measured in log base 10 square feet of living space.  
- A categorical explanatory variable $x_2$: house `condition`, a categorical variable with five levels where `1` indicates “poor” and `5` indicates “excellent.”  

### Multivariate

Let's explore regression models using scatterplots.

```{r}
# Plot interaction model
ggplot(house_prices_reduced, 
       aes(x = log10_size, y = log10_price, col = condition)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(y = "log10 price", 
       x = "log10 size", 
       title = "House prices in Seattle")
```


```{r}
# Plot parallel slopes model
ggplot(house_prices_reduced, 
       aes(x = log10_size, y = log10_price, col = condition)) +
  geom_point(alpha = 0.05) +
  geom_parallel_slopes(se = FALSE) +
  labs(y = "log10 price", 
       x = "log10 size", 
       title = "House prices in Seattle")
```

Let's create a faceted plot of the interaction model.

```{r}
ggplot(house_prices_reduced, 
       aes(x = log10_size, y = log10_price, col = condition)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(y = "log10 price", 
       x = "log10 size", 
       title = "House prices in Seattle") +
  facet_wrap(~ condition)
```

### Regression model

Let's build the interaction model.

```{r}
# Fit regression model:
price_interaction <- lm(log10_price ~ log10_size * condition, 
                        data = house_prices_reduced)
```


```{r}
# Get regression table:
get_regression_table(price_interaction)
```

It is not clear that the interaction model is needed. At most, maybe for a house in condition 5. A machine learning class will help with making a better prediction model.

### Predicting 

Let's use the model to make predictions.  Say you’re a realtor and someone calls you asking you how much their home will sell for. They tell you that it’s in `condition` = 5 and is sized 1900 square feet. Let’s use the interaction model we fit to make predictions! We will use the `augment()` function.

```{r}
price_interaction %>%
  augment(newdata=tibble(condition="5",log10_size=log10(1900)))
```

So the predicted price is

```{r}
10^5.724213	
```

## LC 11.1 (Objective 1)


```{r, include=FALSE, purl=FALSE}
chap <- 11
lc <- 0
```

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Repeat the regression modeling in Subsection 11.2.3 and the prediction making you just did on the house of condition 5 and size 1900 square feet in Subsection 11.2.4, but using the parallel slopes model you visualized in Figure 11.6. 

```{r}
# Fit regression model:
price_parallel <- lm(log10_price ~ log10_size + condition,
  data = house_prices_reduced)
```


```{r}
# Get regression table:
get_regression_table(price_parallel)
```

```{r}
price_parallel %>%
  augment(newdata=tibble(condition="5",log10_size=log10(1900))) %>%
  summarize(price=10^.fitted)
```

Using the rounded numbers from the table:

```{r}
10^(2.88 + 0.096 + 0.837 * log10(1900))
```

***


## Documenting software

-   File creation date: `r Sys.Date()`
-   `r R.version.string`
-   `tidyverse` package version: `r packageVersion("tidyverse")`
-   `moderndive` package version: `r packageVersion("moderndive")`
-   `skimr` package version: `r packageVersion("skimr")`
-   `broom` package version: `r packageVersion("broom")`

