---
title: "Math 300 Lesson 11 Notes"
subtitle: "Simple Linear Regression - Continuous Predictor"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Use the `skimr` package to summarize multiple numerical variables in a data frame.

2. Build a scatterplot to describe the relationship between two continuous, numerical variables; use `geom_smooth()` to visualize the best fit line. 

3. Fit a linear regression model between two variables using the `lm()` function and interpret the output. This includes the interpretation of slope and the use of *association* and not *causation*.    

4. Generate a table of observations, fitted values, and residuals from a linear regression object.  

## Reading 

[Chapter 5 - 5.1](https://moderndive.com/5-regression.html)

## Lesson 

Work through the learning checks LC5.1 - LC5.3. Complete the code as necessary. 

+ Regression can be used for explanatory and predictive purposes. It falls on that line between traditional statistics/econometrics and machine learning. In this course we focus on its more traditional use to interpret the relationship between predictors and a response. Math 378 is our machine learning course and expands on linear regression in this framework.   

+ Note the many different terms for `x` and `y` in regression. These names come from different fields. For example, `y` is called the response, dependent variable, outcome, and output. Meanwhile, `x` is called input, predictor, independent variable, and explanatory variable. Also note that in linear regression, `y` is numerical while `x` can be numerical or categorical.

+ We are using new packages. The `tidyverse` package is a wrapper and actually loads `readr`, `dplyr`, `ggplot2`, and `tidyr`.  

+ The interpretation of the slope has the key phrase **average**. For a one unit change in `x`, the average value of `y` changes by the value of the slope. 


### Setup

```{r message=FALSE}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
``` 

*Create the data needed for the exercises.*

```{r}
# Select only the columns needed. Complete the code and remove comment labels
#evals_ch5 <- evals %>%
#  select(ID, __________ , bty_avg, ____________)
```

Let's look at 5 random rows of data.

```{r}
# Complete the code and remove comment labels
#set.seed(1234)
#evals_ch5 %>%
#  sample_n(size = ___________) 
```

### LC 5.1 (Objective 1)

**(LC 5.1)** Refer to the example in section 5.1.1. Conduct a new exploratory data analysis with the same outcome variable $y$ being `score` but with `age` as the new explanatory variable $x$. Remember, this involves three things:

- Looking at the raw data values.
- Computing summary statistics.
- Creating data visualizations.

What can you say about the relationship between age and teaching scores based on this exploration?

**Solution**: 

- Looking at the raw data values: 
```{r}
# Complete the code and remove comment labels
#glimpse(__________)
```

- Computing summary statistics:

```{r}
# Complete the code and remove comment labels
#evals_ch5 %>%
#  select(_________, _______) %>%
#  skim() 
```  



- Bivariate summary:

```{r}
# Complete the code and remove comment labels
#evals_ch5 %>% 
#  get_correlation(formula = _________ ~ _____________)
```


- Creating data visualizations:

```{r message=FALSE}
# Create scatterplot. Complete the code and remove comment labels
#ggplot(__________, aes(x = ________, y = score)) +
#  geom_jitter(alpha=_________) +
#  labs(
#    x = "__________", y = "Teaching Score",
#    title = "______________________________") +
#  geom_smooth(method = "lm", se = ___________) +
#  theme_classic()
```



### LC 5.2 (Objective 2)

**(LC 5.2)** Fit a new simple linear regression using `lm(score ~ age, data = evals_ch5)` where `age` is the new explanatory variable `x`. Get information about the "best-fitting" line from the regression table by applying the `get_regression_table()` function. How do the regression results match up with the results from your earlier exploratory data analysis?

**Solution**:  

```{r}
# Complete the code and remove comment labels
# Fit regression model:
#score_age_model <- lm(__________ ~ __________, data = ____________)
# Get regression table:
#get_regression_table(score_age_model)
```  



### LC 5.3 (Objective 3)

**(LC 5.3)** Generate a data frame of the residuals of the model where you used `age` as the explanatory $x$ variable.

**Solution**: 

```{r eval=FALSE}

``` 



## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `skimr` package version: `r packageVersion("skimr")`
  * `gapminder` package version: `r packageVersion("gapminder")`  
  * `moderndive` package version: `r packageVersion("moderndive")`
