---
title: "Math 300 Lesson 16 Notes"
subtitle: "Multiple Regression - Related Topics"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

[Chapter 6.4](https://moderndive.com/6-multiple-regression.html#mult-reg-conclusion)

### Setup

```{r message=FALSE}
library(tidyverse)
library(moderndive)
library(skimr)
library(ISLR)
``` 

## Lesson 

Time to makeup missed topics and review for exam.

+ Review the objectives.

+ Rework the learning checks this time without scaffolded code.

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `skimr` package version: `r packageVersion("skimr")`
  * `palmerpenguins` package version: `r packageVersion("palmerpenguins")`  
  * `moderndive` package version: `r packageVersion("moderndive")`
