---
title: "Math 300 Lesson 12 Notes"
subtitle: "Simple Linear Regression - Discrete Predictor"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Explore the relationship between 2 variables, one numerical and one categorical, using summary statistics and visualizations in `R`.  

2. Fit a linear regression model to two variables, one numerical response and one categorical predictor, using the `lm()` function and interpret the output. This includes the interpretation of baseline mean and offsets. 

3. Generate a table of observations, fitted values, and residuals from a linear regression object.  

## Reading 

[Chapter 5.2](https://moderndive.com/5-regression.html#model2)

## Lesson 

Work through the learning checks LC5.4 - LC5.7. Complete the code as necessary.  

+ The response `y` is the numeric variable. Math 378 discusses cases where the response is categorical. Understanding the regression output here is important. There is no line just a baseline average and offsets from that.   

+ The regression output will still predict the **mean** value of the response variable.  

+ The baseline is an **average** and is the first level of the factor based on alphabetic order.


### Setup

```{r message=FALSE}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
``` 

*Create the data needed for the exercises.*

```{r}
# Complete the code and remove comment labels
#gapminder2007 <- _________ %>%
#  filter(year == _______) %>%
#  select(country, lifeExp, ________, gdpPercap)
```

Let's look at 5 random rows of data.

```{r eval=FALSE}
set.seed(1234)
gapminder2007 %>%
  sample_n(size = 5) 
```

### LC 5.4 (Objective 1)

**(LC 5.4)** Conduct a new exploratory data analysis with the same explanatory variable $x$ being `continent` but with `gdpPercap` as the new outcome variable $y$. Remember, this involves three things:

- Most crucially: Looking at the raw data values.
- Computing summary statistics, such as means, medians, and interquartile ranges.
- Creating data visualizations.

What can you say about the differences in GDP per capita between continents based on this exploration?

**Solution**: 

- Looking at the raw data values: 
```{r}
# Complete the code and remove comment labels
#_______________(gapminder2007)
```

- Computing summary statistics, such as means, medians, and interquartile ranges:


```{r, include=FALSE}
my_skim<-skim_with(numeric = sfl(hist = NULL))
```

```{r eval=FALSE}
gapminder2007 %>%
  select(gdpPercap, continent) %>%
  my_skim()
```

- Creating data visualizations:

Create boxplots

```{r}
# Complete the code and remove comment labels
#ggplot(gapminder2007, aes(x = ______________, y = gdpPercap)) +
#  geom_XXXXXX() +
#  labs(
#    x = "Continent", y = "GPD per capita",
#    title = "_________________") +
#  theme_bw()
```


### LC 5.5 (Objective 2)

**(LC 5.5)** Fit a new linear regression using `lm(gdpPercap ~ continent, data = gapminder2007)` where `gdpPercap` is the new outcome variable $y$. Get information about the "best-fitting" line from the regression table by applying the `get_regression_table()` function. How do the regression results match up with the results from your previous exploratory data analysis?

**Solution**: 

```{r}

```


### LC 5.6 (Objective 3)

**(LC 5.6)** Using either the sorting functionality of RStudio's spreadsheet viewer or using the data wrangling tools you learned in Chapter \@ref(wrangling), identify the five countries with the five smallest (most negative) residuals? What do these negative residuals say about their life expectancy relative to their continents?

**Solution**: 

We switched by to life expectancy. We need the model.

```{r}
# Complete the code and remove comment labels
# lifeExp_model <- lm(_________ ~ continent, data = gapminder2007)
```


Use `R`.

```{r}
# Complete the code and remove comment labels
#get_regression_points(lifeExp_model, ID = "______________") %>%
#  arrange(__________) %>%
#  slice_head(n=_______)
```

### LC 5.7 (Objective 3)

**(LC 5.7)** Repeat this process, but identify the five countries with the five largest (most positive) residuals. What do these positive residuals say about their life expectancy relative to their continents?

**Solution**: 

Using `R`.

```{r}



```



## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `skimr` package version: `r packageVersion("skimr")`
  * `gapminder` package version: `r packageVersion("gapminder")`  
  * `moderndive` package version: `r packageVersion("moderndive")`
