
---
title: 'Problem Set 02: Data Wrangling Solutions'
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document: default
  html_document:
    theme: lumen
---

## Documentation: 

We used all the resources available to instructors from the authors of Modern Dive.

### R Packages 

Loading the following packages for this problem set:
```{r warning = F, message = F}
library(ggplot2)
library(dplyr)
```

### The data
Run the following to load and take a `glimpse` of the data:
```{r eval = F}
data(txhousing)
glimpse(txhousing)
```

These data are about housing in Texas. Each row is monthly data for a given city in Texas in a given year. There are multiple years of data for each city. 

***

## Exercise 1

After running all the code above in the console, take a look at the data using `str()`, `glimpse()`, or `View()`. In your report include that last 10 lines of the data file. 

```{r}
tail(txhousing,n=10)
```

***

## Exercise 2
Take a look at the variable descriptions by typing ?txhousing into the **console**. What is the `inventory` variable in this data set?

**Answer: The number of months of inventory; the amount of time it would take to sell all current listings at current pace of sales.**

***

## Exercise 3
Write a code chunk to remove the inventory variable. Save the results in a data frame called `txhousing`. Confirm in the data viewer that the variable has been removed. 

```{r }
txhousing <- txhousing %>% 
  select(-inventory)
```

***

## Exercise 4
Make a data set called `midland_sub` that includes data only from the city of Midland in 2007 & 2014. Print the first 3 rows of `midland_sub`.

```{r }
midland_sub <- txhousing %>% filter(city == "Midland", year == 2007 | year == 2014)
```

```{r}
head(midland_sub,n=3)
```

***

## Exercise 5
Add a column **to the `midland_sub` data set** called `prct_sold` that calculates the percentage of `listings` that were `sold` (`sales/listings * 100`). Be sure to **save** the results also as a data frame called `midland_sub`. Print the first 3 rows of `midland_sub`.

```{r }
midland_sub <- midland_sub %>% 
  mutate(prct_sold = sales/listings *100) 
```

```{r}
head(midland_sub,n=2)
```


***

## Exercise 6
Calculate the **median** percentage of listings that were sold in Midland **in each month of the year** based on your `midland_sub` data set. Save the results of the calculation in an data frame called `midland_summary`. 


```{r }
midland_summary <- midland_sub %>% 
  filter(year==2014) %>%
  group_by(month) %>%
  summarize(median_prct_sold = median(prct_sold))
```

```{r}
midland_summary
```

***

## Exercise 7
Arrange the `midland_summary` in `desc`ending order based on the average percentage of listings, so you can see **which month** had the greatest **average** percentage of listings sold.  You do not need to save the results. 

```{r }
midland_summary  %>% 
  arrange(desc(median_prct_sold))
```

***


## Exercise 8
In August of 2010, what city had the fewest houses listed for sale? (show code and text; do not print out all 46 rows of the data frame.)


**Answer: San Marcos**

```{r eval = F}
txhousing %>% 
  filter(year == 2010, month == 8) %>%
  arrange(listings) %>%
  head()
```

***

## Exercise 9
In 2013, in which month were the most houses sold in Texas? (show code and text)

**Answer: August**


```{r }
txhousing %>% 
  filter(year == 2013) %>%
  group_by(month) %>%
  summarize(top_sales = max(sales)) %>%
  arrange(desc(top_sales))
```


***

## Exercise 10

Generate a single table that shows the total number of houses sold in **Galveston** in **2009 and 2010** (total over the entire period), & the total number of houses sold in **Amarillo** in **2009 and 2010** (total over the entire period). This calculation requires a number of steps, so it might help you to first write out on paper the different steps you will need to take. That will help you set out a "blueprint" for tackling the problem.


```{r }
txhousing %>% 
  filter(city == "Galveston"| city == "Amarillo") %>%
  filter(year == 2009 | year == 2010) %>%
  group_by(city) %>%
  summarize(total_sales = sum(sales))
```

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `dplyr` package version: `r packageVersion("dplyr")`
  * `ggplot2` package version: `r packageVersion("ggplot2")`

