---
title: "Math 300 NTI Lesson 6"
subtitle: "group_by, mutate, and arrange"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Use the `group_by()` function to create aggregated data frames to use with other functions, in particular `summarize()`, to explore, explain, and visualize. 

2. Use the `mutate()` function to create new variables in a data frame to use to explore, explain, and visualize.

3. Use the `arrange()` function to sort data frames to explore, explain, and visualize. 

## Reading 

[Chapter 3.4 - 3.6](https://moderndive.com/3-wrangling.html#groupby)

## Lesson 

Remember that you will be running this more like a lab than a lecture. You want them using `R` and answering questions. Have them open the notes rmd and work through it together.

Work through the learning checks LC3.5 - LC3.12.

+ We changed the scaffolded code method. We have `eval=FALSE` so that `R` does not try to evaluate the code chunk. They have to remove this and then complete the code.

+ It is important to note that the `group_by()` function doesn’t change data frames by itself. Rather it changes the meta-data, or data about the data, specifically the grouping structure. It is only after we apply the `summarize()` function that the data frame changes. The book does a good job explaining meta-data.

+ The `group_by()` can be used on more than two variables but they must be in the same call to `group_by()`. 

+ Using `arrange()` is straight forward expect the use of `desc()` within the `arrange()` call to sort in decreasing order. 

+ As a rough rule of thumb, as long as you are not losing original information that you might need later, it’s acceptable practice to overwrite existing data frames with updated ones.

+ LC 3.6 is difficult. Let them explore and wrestle with this question. The warning can be ignored. We will not experiment with the `.groups` option.

+ LC 3.12 is more difficult as we combined code. The default use of `geom_boxplot()` works for exploring data but we provided code on how to clean up the x-axis. Discuss this code if you want and have time. 

+ The use of `kable()` is only to have the output printed in a form that looks good. This is not something we need to present to the students. 


### Setup  

```{r, include=FALSE, purl=FALSE}
chap <- 3
lc <- 4
```

```{r message=FALSE}
library(nycflights13)
library(ggplot2)
library(dplyr)
```


### LC 3.5 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Recall from Chapter 2 when we looked at plots of temperatures by months in NYC. What does the standard deviation column in the `summary_monthly_temp` data frame, which we need to create from the code at the section 3.4, tell us about temperatures in New York City throughout the year?

**Solution**: 

```{r}
# Code from the book
summary_temp_by_month <- weather %>%
  group_by(month) %>%
  summarize(
    mean = mean(temp, na.rm = TRUE),
    std_dev = sd(temp, na.rm = TRUE)
  )
```


```{r}
# Output
summary_temp_by_month
```

The standard deviation is a quantification of **spread** and **variability**. We
see that the period in November, December, and January has the most variation in
weather, so you can expect very different temperatures on different days in those months.

### LC 3.6 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** What code would be required to get the mean and standard deviation temperature for each day in 2013 for NYC?

**Solution**: 

```{r}
summary(weather)
```

There is only one year 2013 so we don't need to group by it, we could but it would not change anything. 

```{r}
summary_temp_by_day <- weather %>%
  group_by(month, day) %>%
  summarize(
    mean = mean(temp, na.rm = TRUE),
    std_dev = sd(temp, na.rm = TRUE)
  )
```


```{r}
head(summary_temp_by_day)
```


Note: `group_by(day)` is not enough, because `day` is a value between 1-31. We need to `group_by(year, month, day)` or `group_by(month, day)`.

### LC 3.7 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Recreate `by_monthly_origin`, but instead of grouping via `group_by(origin, month)`, group variables in a different order `group_by(month, origin)`. What differs in the resulting dataset?

**Solution**: 

```{r}
by_origin_monthly <- flights %>% 
  group_by(origin, month) %>% 
  summarize(count = n())
```


```{r}
head(by_origin_monthly)
```





```{r}
by_monthly_origin <- flights %>%
  group_by(month, origin) %>%
  summarize(count = n()) 
```

```{r}
head(by_monthly_origin)
```

In `by_monthly_origin` the `month` column is now first and the rows are sorted by `month` instead of origin. If you compare the values of `count` in `by_origin_monthly` and `by_monthly_origin` using the `View()` function, you'll see that the values are actually the same, just presented in a different order. 

### LC 3.8 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** How could we identify how many flights left each of the three airports for each `carrier`?

**Solution**: We could summarize the count from each airport and carrier using the `n()` function, which *counts rows*.

```{r, purl=FALSE,warning=FALSE,message=FALSE}
count_flights_by_airport <- flights %>%
  group_by(origin, carrier) %>%
  summarize(count = n())
```

```{r}
head(count_flights_by_airport,n=10)
```

Note: the `n()` function counts rows, whereas the `sum(VARIABLE_NAME)` function sums all values of a certain numerical variable `VARIABLE_NAME`.

### LC 3.9 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** How does the `filter` operation differ from a `group_by` followed by a `summarize`?

**Solution**: 

* `filter` picks out rows from the original dataset without modifying them, whereas
* `group_by %>% summarize` computes summaries of numerical variables, and hence
reports new values.

### LC 3.10 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** What do positive values of the `gain` variable in `flights` correspond to?  What about negative values?  And what about a zero value?

**Solution**: 

* Say a flight departed 20 minutes late, i.e. `dep_delay = 20`
* Then arrived 10 minutes late, i.e. `arr_delay = 10`.
* Then `gain = dep_delay - arr_delay = 20 - 10  = 10` is positive, so it "made up/gained time in the air."
* 0 means the departure and arrival delay times were the same, so no time was made up in  the air. We see in most cases that the `gain` is near 0 minutes.


### LC 3.11 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Could we create the `dep_delay` and `arr_delay` columns by simply subtracting `dep_time` from `sched_dep_time` and similarly for arrivals?  Try the code out and explain any differences between the result and what actually appears in `flights`.

**Solution**: No because you can't do direct arithmetic on times. The difference in time between 12:03 and 11:59 is 4 minutes, but `1203-1159 = 44`. Plus there are time zones, departure and arrival times are in the local timezone, which cause problems with simple subtraction.

```{r purl=FALSE}
LC3.11<- flights %>%
  mutate(time_gain=dep_time-arr_time,gain = dep_delay - arr_delay) %>%
  select(air_time,dep_time,arr_time,time_gain,dep_delay,arr_delay, gain)
```

```{r}
head(LC3.11)
```


### LC 3.12 (Objective 2)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** What can we say about the distribution of `gain`?  Describe it in a few sentences using a boxplot and the `gain_summary` data frame values.

**Solution**: We must create the data frame from the notes. We copied the code from the book, we had to copy and combine two chunks of code.

```{r}
gain_summary <- flights %>%
  mutate(gain = dep_delay - arr_delay) %>%
  summarize(
    min = min(gain, na.rm = TRUE),
    q1 = quantile(gain, 0.25, na.rm = TRUE),
    median = quantile(gain, 0.5, na.rm = TRUE),
    q3 = quantile(gain, 0.75, na.rm = TRUE),
    max = max(gain, na.rm = TRUE),
    mean = mean(gain, na.rm = TRUE),
    sd = sd(gain, na.rm = TRUE),
    missing = sum(is.na(gain))
  )
```

```{r}
gain_summary
```

```{r warning=FALSE}
flights %>%
  mutate(gain = dep_delay - arr_delay) %>%
  ggplot(aes(x=1,y=gain)) +
  geom_boxplot() +
  scale_x_continuous(breaks = NULL) +
  theme(axis.title.x = element_blank()) +
  theme_classic()
```


Most of the time the gain is a little above zero (the median is 7, meaning gain is above 0 at least 50% of the time) and between -50 and 50 minutes. There are some extreme cases however!

## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `ggplot2` package version: `r packageVersion("ggplot2")`
  * `dplyr` package version: `r packageVersion("dplyr")`
  * `nycflights13` package version: `r packageVersion("nycflights13")`
