---
title: "Math 300 NTI Lesson 5"
subtitle: "filter and summarize"
author: "Professor Bradley Warner"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    css: NTI.css
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Use the `filter()` function and logical operators to subset a data frame. 

2. Use the `summarize()` function with appropriate `R` functions to summarize variables in a data frame. 

3. Explain the possible impacts of simply ignoring missing values.


## Reading 

[Chapter 3 - 3.3](https://moderndive.com/3-wrangling.html)

## Lesson 

Remember that you will be running this more like a lab than a lecture. You want them using `R` and answering questions. Have them open the notes rmd and work through it together.

Work through the learning checks LC3.1 - LC3.4.

+ As you have probably heard: about 80% of your work analyzing data will be data acquisition and wrangling. 

+ The order of logical operations in the `filter()` function work from left to right. Parentheses, however, are executed first. Be careful and when in doubt use parentheses. 

+ Pay attention to functions that default an `NA` for variables with a missing value. Functions such as `mean()` and `sd()` are examples.

+ This [pdf](https://moderndive.github.io/moderndive_labs/static/PS/PS03_pre_lecture.pdf) summarizes the pipe operator and the main data wrangling functions. There should be time to discuss it in this lesson.

### Setup  

```{r, include=FALSE, purl=FALSE}
chap <- 3
lc <- 0
```

```{r message=FALSE}
library(nycflights13)
library(ggplot2)
library(dplyr)
```

*Notice that order of logical operators doesn't matter in this example but the parentheses help the reading of the code.*

```{r}
flights %>% 
  filter(origin == "JFK", (dest == "BTV" | dest == "SEA"), month >= 10) %>%
  glimpse()
```


```{r}
flights %>% 
  filter(origin == "JFK", dest == "BTV" | dest == "SEA", month >= 10) %>%
  glimpse()
```

### LC 3.1 (Objective 1)

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** What's another way using the "not" operator `!` to filter only the rows that are not going to Burlington, VT nor Seattle, WA in the `flights` data frame? Test this out using the code above.

**Solution**: 

```{r, eval=FALSE, purl=FALSE}
# Original in book
not_BTV_SEA <- flights %>%
  filter(!(dest == "BTV" | dest == "SEA"))

# Alternative way
not_BTV_SEA <- flights %>%
  filter(!dest == "BTV" & !dest == "SEA")

# Yet another way
not_BTV_SEA <- flights %>%
  filter(dest != "BTV" & dest != "SEA")
```

### LC 3.2 (Objective 3) 


**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Say a doctor is studying the effect of smoking on lung cancer for a large number of patients who have records measured at five year intervals. She notices that a large number of patients have missing data points because the patient has died, so she chooses to ignore these patients in her analysis. What is wrong with this doctor's approach?

**Solution**: The missing patients may have died of lung cancer! So to ignore them might seriously **bias** your results! It is very important to think of what the consequences on your analysis are of ignoring missing data! Ask yourself:

* There is a systematic reasons why certain values are missing? If so, you might be biasing your results!
* If there isn't, then it might be ok to "sweep missing values under the rug."

### LC 3.3 (Objective 2) 

```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
```


```{r}
summary_temp
```


**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Modify the above `summarize` function to create `summary_temp` to also use the `n()` summary function: `summarize(count = n())`. What does the returned value correspond to?

**Solution**: It corresponds to a count of the number of observations/rows:

```{r, purl=FALSE}
weather %>%
  summarize(count = n())
```

### LC 3.4 (Objective 2) 

**`r paste0("(LC", chap, ".", (lc <- lc + 1), ")")`** Why doesn't the following code work?  Run the code line by line instead of all at once, and then look at the data.  In other words, run `summary_temp <- weather %>% summarize(mean = mean(temp, na.rm = TRUE))` first.

```{r eval=FALSE}
summary_temp <- weather %>%
  summarize(mean = mean(temp, na.rm = TRUE)) %>%
  summarize(std_dev = sd(temp, na.rm = TRUE))
```

**Solution**: Consider the output of only running the first two lines:

```{r}
weather %>%
  summarize(mean = mean(temp, na.rm = TRUE))
```

Because after the first `summarize()`, the variable `temp` disappears as it has been collapsed to the value `mean`. So when we try to run the second `summarize()`, it can't find the variable `temp` to compute the standard deviation of. 



## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `ggplot2` package version: `r packageVersion("ggplot2")`
  * `dplyr` package version: `r packageVersion("dplyr")`
  * `nycflights13` package version: `r packageVersion("nycflights13")`
