---
title: "Math 300 Lesson 6 Notes"
subtitle: "group_by, mutate, and arrange"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Use the `group_by()` function to create aggregated data frames to use with other functions, in particular `summarize()`, to explore, explain, and visualize. 

2. Use the `mutate()` function to create new variables in a data frame to use to explore, explain, and visualize.

3. Use the `arrange()` function to sort data frames to explore, explain, and visualize. 

## Reading 

[Chapter 3.4 - 3.5](https://moderndive.com/3-wrangling.html#groupby)

## Lesson 

Work through the learning checks LC3.5 - LC3.12. Complete code as necessary.

+ It is important to note that the `group_by()` function doesn’t change data frames by itself. Rather it changes the meta-data, or data about the data, specifically the grouping structure. It is only after we apply the `summarize()` function that the data frame changes. The book does a good job explaining meta-data.

+ The `group_by()` can be used on more than two variables but they must be in the same call to `group_by()`. 

+ Using `arrange()` is straight forward expect the use of `desc()` within the `arrange()` call to sort in decreasing order. 

+ As a rough rule of thumb, as long as you are not losing original information that you might need later, it’s acceptable practice to overwrite existing data frames with updated ones.

### Setup  

```{r message=FALSE}
library(nycflights13)
library(ggplot2)
library(dplyr)
```


### LC 3.5 (Objective 1)

**(LC 3.5)** Recall from Chapter 2 when we looked at plots of temperatures by months in NYC. What does the standard deviation column in the `summary_monthly_temp` data frame, which we need to create from the code at the section 3.4, tell us about temperatures in New York City throughout the year?

**Solution**: 

```{r}
# Complete the code and then remove comment symbols
# summary_temp_by_month <- _________ %>%
#   group_by(______) %>%
#   summarize(
#     mean = mean(_______, na.rm = TRUE),
#     std_dev = sd(_______, na.rm = TRUE)
#   )
```




### LC 3.6 (Objective 1)

**(LC 3.6)** What code would be required to get the mean and standard deviation temperature for each day in 2013 for NYC?

**Solution**: 

```{r}
#Complete the code and then remove comment symbols
# summary_temp_by_day <- weather %>%
#   group_by(________, ________) %>%
#   summarize(
#     mean = mean(temp, na.rm = TRUE),
#     std_dev = sd(temp, na.rm = TRUE)
#   )
```


```{r}
#head(summary_temp_by_day)
```



### LC 3.7 (Objective 1)

**(LC 3.7)** Recreate `by_monthly_origin`, but instead of grouping via `group_by(origin, month)`, group variables in a different order `group_by(month, origin)`. What differs in the resulting dataset?

**Solution**: 

```{r eval=FALSE}
# Complete the code and then remove comment symbols
# by_origin_monthly <- flights %>% 
#   group_by(________, __________) %>% 
#   summarize(count = ______)
```

```{r eval=FALSE}
# Complete the code and then remove comment symbols
# by_monthly_origin <- flights %>%
#   group_by(________, __________) %>% 
#   summarize(count = ______) 
```



### LC 3.8 (Objective 1)

**(LC 3.8)** How could we identify how many flights left each of the three airports for each `carrier`?

**Solution**: 

```{r}
# Complete the code and then remove comment symbols
# count_flights_by_airport <- flights %>%
#   group_by(_________, _________) %>%
#   summarize(count = _______)
```

```{r}
# Complete the code and then remove comment symbols
#head(count_flights_by_airport,n=______)
```


### LC 3.9 (Objective 1)

**(LC 3.9)** How does the `filter` operation differ from a `group_by` followed by a `summarize`?

**Solution**: 






### LC 3.10 (Objective 2)

**(LC 3.10)** What do positive values of the `gain` variable in `flights` correspond to?  What about negative values?  And what about a zero value?

**Solution**: 






### LC 3.11 (Objective 2)

**(LC 3.11)** Could we create the `dep_delay` and `arr_delay` columns by simply subtracting `dep_time` from `sched_dep_time` and similarly for arrivals?  Try the code out and explain any differences between the result and what actually appears in `flights`.

**Solution**: 

```{r}
# Complete the code and then remove the comment symbols 
# LC3.11<- flights %>%
#   mutate(time_gain=dep_time-arr_time, gain = _____________) %>%
#   select(air_time,dep_time,arr_time,time_gain,dep_delay,arr_delay, gain)
```


### LC 3.12 (Objective 2)

**(LC 3.12)** What can we say about the distribution of `gain`?  Describe it in a few sentences using a boxplot and the `gain_summary` data frame values.

**Solution**: We must create the data frame from the notes. We have to copy and combine two chunks of code.

```{r}
# Complete the code and then remove comment symbols
# gain_summary <- ______ %>%
#   mutate(gain = _______ - __________) %>%
#   summarize(
#     min = min(_________, na.rm = TRUE),
#     q1 = quantile(gain, 0.25, na.rm = TRUE),
#     median = quantile(gain, 0.5, na.rm = TRUE),
#     q3 = quantile(gain, 0.75, na.rm = TRUE),
#     max = max(gain, na.rm = TRUE),
#     mean = _________(gain, na.rm = TRUE),
#     sd = _________(gain, na.rm = TRUE),
#     missing = sum(is.na(___________))
#   )
```


```{r eval=FALSE}
# Complete the code and then remove the comment symbols
# flights %>%
#   mutate(gain = _____________ - ____________) %>%
#   ggplot(aes(x=1,y=gain)) +
#   ___________() +
#  theme_classic()
```



## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `ggplot2` package version: `r packageVersion("ggplot2")`
  * `dplyr` package version: `r packageVersion("dplyr")`
  * `nycflights13` package version: `r packageVersion("nycflights13")`
