---
title: "Math 300 Lesson 8 Notes"
subtitle: "Importing Data"
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%B, %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Import csv and Excel data files into `R`.  

2. Explain and use appropriately the concept of tidy data. 

3. Create a tidy data frame using the appropriate functions in `R`. 

## Reading 

[Chapter 4 - 4.2](https://moderndive.com/4-tidy.html)

## Lesson 

Work through the learning checks LC4.1 - LC4.3. Complete the code as necessary.  

+ Although this chapter seems straightforward, it is not. Thinking about the form you want the data in means defining the observational unit. 

+ From Wickham's paper: Tidy data is a standard way of mapping the meaning of a dataset to its structure. A dataset is messy or tidy depending on how rows, columns and tables are matched up with observations, variables and types. You must think about your data before analyzing it. 

+ The book makes it seem that long data frames are tidy and wide ones are not. Be careful, this can be too simplistic. 

+ The `pivot_longer()` function is difficult when you first use it. The function arguments can be confusing. The `names_to` and `values_to` are really just asking for the names of columns when done. The `names_to` takes the column names and creates a variable with the assigned name. The `values_to` takes the values in the selected columns and makes them a variable. The `cols` is subtle and can be done in a variety of ways. Practice, run `?pivot_longer` or go to the [tidyverse](https://tidyr.tidyverse.org/dev/articles/pivot.html#pew) for more examples.  

### Setup

```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(nycflights13)
library(fivethirtyeight)
``` 

### Import data 4.1.2 (Objective 1)

Repeat the import of `dem_score.xlsx` into `R`. Experiment with the options in the GUI. Also import `https://moderndive.com/data/dem_score.csv` using the `Import Dataset` icon under the **Environment** tab.

### LC 4.1 (Objective 2)

**(lc 4.1)** What are common characteristics of "tidy" datasets?

**Solution**: 



### LC 4.2 (Objective 2)

**(LC 4.2)** What makes "tidy" datasets useful for organizing data?

**Solution**: 



### LC 4.3 (Objective 2)

**(LC 4.3)** Take a look the `airline_safety` data frame included in the `fivethirtyeight` data. Run the following:

```{r, eval=FALSE}
head(airline_safety)
```

After reading the help file by running `?airline_safety`, we see that `airline_safety` is a data frame containing information on different airlines companies' safety records. This data was originally reported on the data journalism website FiveThirtyEight.com in Nate Silver's article ["Should Travelers Avoid Flying Airlines That Have Had Crashes in the Past?"](https://fivethirtyeight.com/features/should-travelers-avoid-flying-airlines-that-have-had-crashes-in-the-past/). Let's ignore the `incl_reg_subsidiaries` and `avail_seat_km_per_week` variables for simplicity:

```{r}
airline_safety_smaller <- airline_safety %>%
  select(airline, starts_with("fatalities"))
```


```{r}
head(airline_safety_smaller)
```


This data frame is not in "tidy" format. How would you convert this data frame to be in "tidy" format, in particular so that it has a variable `fatalities_years` indicating the incident type/year and a variable `count` of the counts?

**Solution**: 



## Documenting software 

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `ggplot2` package version: `r packageVersion("ggplot2")`
  * `tidyr` package version: `r packageVersion("tidyr")`
  * `readr` package version: `r packageVersion("readr")`  
  * `dplyr` package version: `r packageVersion("dplyr")`
  * `nycflights13` package version: `r packageVersion("nycflights13")`
  * `fivethirtyeight` package version: `r packageVersion("fivethirtyeight")`  